\chapter{Track based feature reduction}

A brief overview of the chapter.

\section{Feature reduction}
I will discuss the need for feature reduction. 
As we are planning for a smartphone, 
the processing power matters a lot. 
The reduced features will definitley 
result in better efficiency.
 
\subsection{Related Work}
I will discuss some of the related work 
done to reduce the features.

I will mention the research works from 
which our work is inspired from.


\section{Track based feature reduction}

\subsection{Average descriptors}
I will discuss the strength of average descriptors 
by discussing the distances computed between 
different similar features.

\subsection{Tracks Generation}
I will discuss the ways used for the track generation.
I will state the algorithm used for track generation 
towards the end.
 
\subsection{Reduced Feature Set}
I will discuss different type of reduced features used 
in the work.

\section{Scene Representation}
Every scene can be represented by a 
single feature file. I will discuss thee need 
to split the feature file 
into more than one file for visual BoW.
I will also narrate the algorithm used by me.


\section{Data sets}
\begin{enumerate}
\item David Nister
\item Owheo
\item Pasadena
\item Commerce
\end{enumerate}
\section{Performance Metrics}
\section{Results}
\subsection{Feature reduction statistics}
\subsection{Naive Matching}
\subsection{Visual BoW}
\subsection{Scene Clustering}
\subsection{Scene Confusion Matrix}
\section{Conclusion}